Representing and Reasoning About Spatial Regions Defined by Context

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
JK: qsr-extraction.tex updated, spatial relation images added, bib file updated

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing spatial regions, these approaches do not consider the semantic information necessary for identifying \emph{context-dependent} regions. While steps have been taken towards incorporating this knowledge  into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, systems usually only refer to the categories of whole rooms, \cite{Hawes/etal:2010b,Vasudevan/etal:2008,Galindo/etal:2005}, whilst others are able to temporarily mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about CDSRs by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.  The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics.

>>>>

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/target-success.png}
  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region. The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
  \label{fig:success}
\end{figure}

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions. While steps have been taken toward incorporating this knowledge  (i.e. the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
 % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.

There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.  The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
JK: qsr-extraction.tex updated, spatial relation images added, bib file updated

\from{Nick}{Do these 8 relations have names? We need them here so we can refer to the predicates later.}

For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}. The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object. The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.

We generate these relations as follows. Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above. This angle is computed by taking the dot product of these two vectors. $\dots$

%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
%%Desk 1 and Desk 2 are not labeled in the Figure.

%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}. Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint. To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}. The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate. Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}. These qualitative spatial relationships provide the structure necessary for analogical processing.

>>>>

For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}. The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms, \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}. The model accommodates both direction and distance as factors in the relative position of objects.

The relations we compute between each given \emph{landmark} object and its adjacent neighbours are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object. The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$; and, the predicates used to denote these relations are named accordingly, e.g. \textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.

We generate the strengths of these spatial relations as follows. First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects in the room. Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the vector from the origin (the landmark's centroid) to the neighbour's location and the direction vector of the relation. This angle is computed by taking the dot product of these two vectors. These two spatial components, distance and angular deviation, are then integrated to compute the strength of a given relationships as follows:

\begin{equation}
strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
\end{equation}

Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region. These spatial relationships between adjacent objects provide the structure necessary for analogical processing.

\begin{figure}
\centerline{
\includegraphics[width=2in]{./images/spatialRelations/xOneYZeroWhite.png}
}
\caption{A visualisation of a the strength of a spatial relation across a region. In this figure the landmark is located at the center of the image and the direction vector used was $<1,0>$ (i.e. right of the landmark). The darker the pixel the stronger the spatial relation is deemed to be at that point.}
\label{fig:spr}
\end{figure}
